Risk and Trust in Open Systems:\\ \Large Towards formalising Permission and Authority, and specifying policies for Object Capability Patterns

### 2017-09-08 shupeng_msc_report.tex
updated code

\makeatletter
\DeclareRobustCommand{\emp}{%
  \@nomath\em \if b\expandafter\@car\f@series\@nil
  \normalfont \else \sffamily \bfseries \fi}
\makeatother
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al. \cite{drossopoulou2016}, and we contribute to the literature by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper). We show how our definitions can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code. Another contribution we made is that we propose and show how the concept of domination over objects can be used to specify and reason about safe or vulnerable cooperation, which is novel but inspired by the work of Clarke et al. on ownership types \cite{clarke1998}. Our paper then show how using a mix of our proposed formalisations of permission, authority, domination, and Hoare logics, we can formally specify well-established Object-Capability (OCap) patterns in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.  In particular, we reason about protecting the properties of nodes in a DOM Tree, inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within the capability-safe language Pony \cite{clebsch2015}. We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system. Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system, using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
}
\begin{document}
\let\sf\textsf
\input{titlepage}
\raggedbottom
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}
\clearpage
\thispagestyle{empty}
\begin{minipage}{\textwidth}
\begin{center}
\textit{Acknowledgements}
\end{center}
\small
I am immensely grateful to my supervisor Prof. Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from. I will miss the countless hours we have worked together on this thesis.\\

@@

It is therefore not surprising that \textit{thinking} about security while building a complex system of cooperation that involves the interplay between trusted(known) and untrusted(unknown) code components is extremely challenging to any programmer. To some programmers, security, is often treated as a \textit{separate} mental burden to be considered separately from the functionality of a system. In recent years the Object-Capability (OCap) model has received increasing attention as a compelling approach to building robust, distributed systems that promote what Miller\cite{miller2006} calls \textit{cooperation without vulnerability}. The OCap model attemps to address these two issues by alleviating security as a separate concern from the mind of the programmer. This is done through leveraging the object-oriented programming paradigm and imposing certain prohibitions or restrictions. Within the OCap literature, there are two central concepts that describe how objects interact: permission and authority. The OCap literature also features prominently how using an OCap system would accomplish cooperation without vulnerability through Object-Capability Patterns (OCap patterns). OCap patterns are important, because they demonstrate how objects can cooperate with unknown code while preserving desirable properties in a system.\\
\subsection{Our contributions}
While there have been considerable work done on OCap patterns in programming languages \cite{miller2003b,miller2006,murray2010}, less work has been done on formally specifying them. The only works we know so far are \cite{drossopoulou2015b,devriese2016,swasey2017}, which are all very recent in the literature. Our paper builds on the works of \cite{drossopoulou2015b}, because we claim that the methodology of using OCap policies to specify OCap patterns to be intuitive and straightforward. While they may seem less sophisticated than the methodologies in other works, we believe they are powerful enough to reason convincingly about the interactions between trusted code and code of unknown provenance. To formally specify OCap patterns in the style of OCap policies, our paper contributes to the literature by proposing formal predicates for permission and authority that describe access dynamics between objects. Miller first proposed the concepts of permission and authority in his PhD thesis\cite{miller2006}, following which the formal definitions of these concepts were proposed and refined in the work of Drossopoulou et al. \cite{drossopoulou2016}. Our paper defines permission and authority in a formal way that is different from the aforementioned works, but we generally adopt the same informal meanings (that having permission means the right to invoke an object's behaviours, and authority to mean the ability to cause effects). More precisely, our paper defines \textit{permission} as having the reference of another object (right to invoke that object's behaviours), either directly or indirectly---through multiple object reference(s)---and either currently or eventually---in some current or future state of the system; and defines \textit{authority} to mean being \textit{able} to invoke an object's behaviours; i.e. we distinguish between having the \textit{right to do} something (permission) and \textit{being able to do} something (authority)\footnote{An anology would be having the right to perform an impossible task does not imply the task becomes possible.}. Furthermore, our paper implicitly treats invoking behaviours of objects as 'effects' in the system. We claim that these insights are novel and our motivation for defining them can be found in \cref{sec:permauth}.\\

To put our formal definitions of permission and authority in practice and demonstrate their usefulness, we further propose the novel use of the concept of domination over objects that is inspired by the works of \cite{clarke1998} on ownership types. Our domination predicate describes a special set of objects that dominate a particular object---all permissions to use the particular object have to involve the permission of a member of the dominating set, and all authorities have to involve the authority of a member of the dominating set.\\

With our formal definitions of permission, authority, and domination, our paper then argues in \cref{sec:ocapreasoning} on how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability} and \textit{protection}. We claim that our framework allows one to understand and reason about these concepts in a highly straightforward and intuitive way without treating security as a separate burden from the functionality of the system, which we hope would encourage programmers to seek the use of the object-capability model to build robust and secure systems.\\

To show this in practice, our paper provides an illustration of three OCap patterns found in \cref{sec:figdom,sec:figcaretaker,sec:figmem}, and we focus only on providing the OCap policy specifications for the DOM Tree pattern in \cref{sec:dompolicies}. We choose the DOM Tree pattern because we are convinced of the importance of reasoning about unknown code using the object-capability framework in the web security space, where guest/unknown code (such as third party scripted advertising or widgets) are pervasive in the web today (details can be found in the work by Maffeis et al.\cite{maffeis2010}). We show how using our OCap policies, we can specify the preservation of properties of nodes, in the presence of unknown code execution, in \cref{sec:treereasoning}. Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years, and we recognise that further work has to be done in the field of smart contracts on the blockchain technology.

\subsection{Overview of Paper}
In our paper, we first give an overview background history on the OCap model in \cref{sec:ocapmodel}. Readers familiar with the object-capability model and its languages are encouraged to skim through this section. In \cref{sec:specs}, we seek to formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other. Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language. That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapreasoning}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily. We further solidify how our methodology can be used to reason convincingly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}. We end off with some preliminary insights on Ethereum in \cref{sec:ethereum}.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{Object-Capability Model}\label{sec:ocapmodel}
\subsection{From Object to Capability}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}. When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.

@@

%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\newpage
\section{Formal specifications}\label{sec:specs}
In this section we describe a small object-oriented language in \cref{sec:speclang} and define predicates that allow us to specify object-capability policies in \cref{sec:permauth}. For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}. To express more precisely how an object can interact with another object, we use the ideas of permission and authority in Miller's\cite{miller2006} and Drossopoulou et al.'s\cite{drossopoulou2016} works. In their papers and our paper, we generally try to capture the informal meaning of:
\begin{itemize}
\item permission: access right of an object to invoke a behaviour of an object
\item authority: ability to cause effects
\end{itemize}
In the approach of our paper however, we contribute to the literature by taking a novel approach in formally defining both permission and authority, though we argue they generally capture the same informal meanings. For permission, we have four different modes permissions specified by four MayAccess predicates. Given a module M and state \losigma,:
\begin{tabbing}
MayAccess$^{Dir,Now}$(o,o$'$): \=o has an object reference of o$'$\\
MayAccess$^{Ind,Now}$(o,o$'$): o contains a reference that points to an object that has a reference\\
\>to another object---thus forming a series of object reference(s)---\\ \> where the last object in the series has an object reference of o$'$\\
MayAccess$^{Dir,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(o,o$'$) holds\\
MayAccess$^{Ind,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(o,o$'$) holds\\
\end{tabbing}
For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}. While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker ---we mean the ability to invoke a method of an object. Therefore implicitly, we consider method invocation as an effect in the system. Given a module M and state \losigma,:
\begin{tabbing}
MayCall(o,o$'$): \=there exist a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
\end{tabbing}

We further elaborate these predicates in \cref{MayAccess} and \cref{MayCall}.
\subsection{Specification Language}\label{sec:speclang}
In our small specification language, we briefly describe the defintions of module, class, runtime state, interpretation, execution, arising configurations, and assertion.
\textbf{Module:}\\
A module maps class identifiers to class descriptions:\\
\begin{logic}
M \loin\ Module = ClassId \locup\ SpecId\\\ablock\qquad\qquad\quad $\rightarrow$ \\
\ablock \qquad \qquad \quad (ClassDescr \locup\ Specification)
\end{logic}\\

\textbf{Class:}\\
Class definitions describe how a class contributes to the runtime behaviour of a program for which we use its field and method declarations. We define method bodies as consisting a sequences of statements; these statements can be field read or internal field assignments (only allowed if the object is \texttt{this}), conditionals, and method calls. We also define all classes to be private which mean only objects belonging to class C may construct objects of class C.\\

@@

 \begin{logic}
\losigma\ \loin\ state \quad\= = frame \lotimes\ heap \\
\lophi \loin\ frame \>= StackId $\rightarrow$ val \\
\lochi \loin\ heap \>= addr $\rightarrow$ object \\
v \loin\ val \>= \{\texttt{null}, \texttt{true}, \texttt{false}\} \locup\ \textit{addr} \locup\ $\mathbb{N}$ \\
StackId \>= \{\textit{this}\} \locup\ VarId \locup\ ParId \\
object \>= ClassId \lotimes (FieldId $\rightarrow$ val)\\
$\iota$,$\iota$',.. \>\loin\ \textit{addr}\\
\end{logic}

@@

@@

The set Reach(M,\losigma,stmts) collects all configurations reachable during execution of \losigma,stmts. Note that the function Reach(M,\losigma,stmts) is defined even when execution diverges so it may be an infinite set. We therefore can give meaning to capability policies without requiring termination.\\

@@

We say that a class \sf{C} satisfies an assertion \sf{A} in the context of a module \sf{M}, if in all runtime configuration \losigma$'$ arising from execution of any module \sf{M$'$} linked with module \sf{M}, all objects of class \sf{C} satisfies \sf{A}:\\

@@

M,\losigma\ \loturns\ \=MayAccess$^{Dir,Now}$(x,e) \loiff \\
\>\loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
\loor \\
\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y}{\losigma} = \loexec{e}{\losigma})\\
\\
 M,\losigma\ \loturns\ \=MayAccess$^{Dir,Eve}$(x,e) \loiff \\
\>\loexists \losigma$'$ \loin Arising(M,\losigma).\\
\>M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(x,e)\\
\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(x,e) \loiff \\
\>\loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
\loor \\
\> \loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y.\loconj{f}}{\losigma} = \loexec{e}{\losigma})\\
\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(x,e) \loiff \\
\>\loexists \losigma$'$ \loin Arising(M,\losigma).\\
\>M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(x,e)\\
\hr
\end{logic}

@@

Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions. Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions. \\
\begin{table}[htb]
\caption{MayAccess Relations}\label{tab:accessrelations}
\centering
\begin{tabular*}{0.5\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &\\
\end{tabular*}
\end{table}\\
We surmarise the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}. Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities. MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$). The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).

@@

\begin{logic}
\hr\\
\emp{Definition---\emp{[MayAffect from Drossopoulou et al.\cite{drossopoulou2016}]}}\\
M,\losigma\ \loturns\ \=MayAffect(x,e) if there exists a method \sf{m}, argument(s) \loconj{a}, state \losigma', such that:\\
\ablock M,\losigma\ \=e.m(\loconj{a}) \loleadsto\ \losigma', and\\
\>\loexec{e}{M,\losigma} \loneq\ \loexec{e$'$}{M,\losigma$'$}\\
\hr
\end{logic}

@@

\begin{logic}
\hr\\
\emp{Lemma---[Object as receiver requires method invocation]}\\
M,\losigma\ \loturns\ \=\losigma(this) = \loexec{o}{\losigma}\loimplies \\
\> \loexists \losigma$'$,m,\loconj{a}. [ M,\losigma$'$, o.m(\loconj{a}) \loleadsto\ \losigma\ ]\\
\hr
\end{logic}

We make clear that for an object to be a receiver in a particular state \losigma, we require a method to exist in the object in a prior state \losigma$'$ that upon invocation can lead to state \losigma. This follows from our definition of execution.\\

>>>>

\makeatletter
\DeclareRobustCommand{\emp}{%
  \@nomath\em \if b\expandafter\@car\f@series\@nil
  \normalfont \else \sffamily \bfseries \fi}
\makeatother
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al. \cite{drossopoulou2016}, and we contribute by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper). Another contribution we made is that we propose and show how the concept of domination over objects can be used to reason about OCap policies, which is novel but inspired by the work of Clarke et al. on ownership types \cite{clarke1998}. We show how our definitions above can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code. Our paper then shows how the use of our concepts of permission, authority, and domination within extended Hoare logics can formally specify well-established Object-Capability (OCap) patterns, namely the DOM Tree pattern, Caretaker pattern and Membrane pattern, in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.  In particular, we show how we can reason about protecting the properties of nodes in the DOM Tree pattern, which is inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within our methodology and the capability-safe language Pony \cite{clebsch2015}. We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system. Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system, using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
}
\begin{document}
\let\sf\textsf
\input{titlepage}
\raggedbottom
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}
\clearpage
\thispagestyle{empty}
\begin{minipage}{\textwidth}
\begin{center}
\textit{Acknowledgements}
\end{center}
\small
I am immensely grateful to my supervisor Prof. Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from. I will miss the countless hours we have worked together on this thesis.\\

@@

It is therefore not surprising that \textit{thinking} about security while building a complex system of cooperation that involves the interplay between trusted(known) and untrusted(unknown) code components is extremely challenging to any programmer. To some programmers, security, is often treated as a \textit{separate} mental burden to be considered separately from the functionality of a system. In recent years the Object-Capability (OCap) model has received increasing attention as a compelling approach to building robust, distributed systems that promote what Miller\cite{miller2006} calls \textit{cooperation without vulnerability}. The OCap model addresses these two issues by alleviating security as a separate concern from the mind of the programmer. This is done through leveraging the object-oriented programming paradigm and imposing certain prohibitions or restrictions. Within the OCap literature, there are two central concepts that describe how objects interact: permission and authority. The OCap literature also features prominently how using an OCap system would accomplish cooperation without vulnerability through Object-Capability Patterns (OCap patterns). OCap patterns are important, because they demonstrate how objects can cooperate with unknown code while preserving desirable properties in a system.\\
\subsection{Our contributions}
While there have been considerable work on OCap patterns in programming languages \cite{miller2003b,miller2006,murray2010,drossopoulou2014b}, less work has been done on formally specifying them. The only works we know so far are \cite{drossopoulou2015b,devriese2016,swasey2017}, which are all very recent in the literature. Our paper builds on the works of \cite{drossopoulou2015b}, but we propose a different methodology for specifying OCap patterns which we claim is intuitive and straightforward. While ours may seem less sophisticated than the methodologies in other works, we believe that ours are powerful enough to reason convincingly about the interactions between trusted code and code of unknown provenance.\\

To formally specify OCap patterns in the style of OCap policies, our paper contributes to the literature by proposing formal predicates for permission and authority that describe dynamic access between objects. Miller first proposed the concepts of permission and authority in his PhD thesis\cite{miller2006}, following which the formal definitions of these concepts were proposed and refined in the work of Drossopoulou et al. \cite{drossopoulou2016}. Our paper defines permission and authority in a formal way that is different from the aforementioned works, but we generally adopt similar intuitions (that having permission means the right to invoke an object's behaviours, and authority to mean the ability to cause effects). More precisely, our paper defines \textit{permission} as having the reference of another object (right to invoke that object's behaviours), either directly or indirectly---through multiple object reference(s)---and either currently or eventually---in some current or future state of the system; and defines \textit{authority} to mean being \textit{able} to invoke an object's behaviours; i.e. we distinguish between having the \textit{right to do} something (permission) and \textit{being able to do} something (authority)\footnote{An analogy would be having the right to perform an impossible task does not imply the task becomes possible.}. Our paper also implicitly treats invoking behaviours of objects as 'effects' in the system. We claim that these insights are novel and our motivation for defining them can be found in \cref{sec:permauth}.\\

We further propose the novel use of the concept of domination over objects that is inspired by the works of \cite{clarke1998} on ownership types. Our domination predicate describes a special set of objects that dominate a particular object---all permissions to use the particular object have to involve the permission of a member of the dominating set, and all authorities have to involve the authority of a member of the dominating set.\\

With our formal definitions of permission, authority, and domination, our paper then argues in \cref{sec:ocapreasoning} on how we can use them to reason and distinguish concepts of \textit{isolation}, \textit{cooperation}, \textit{vulnerability}, \textit{safety} and \textit{protection}. We claim that our framework allows one to understand and reason about these concepts in a highly straightforward and intuitive way without treating security as a separate burden from the functionality of the system, which we hope would encourage programmers to seek the use of the object-capability model to build robust and secure systems.\\

To put our formal definitions of permission, authority, and domination in practice and demonstrate their usefulness, 
our paper provides an illustration of three OCap patterns, DOM Tree pattern, Caretaker pattern, and Membrane pattern, which can be found respectively in \cref{sec:figdom,sec:figcaretaker,sec:figmem}. Our paper also provides the corresponding OCap policies based on our formal definitions for these three patterns. In \cref{sec:treereasoning}, we choose the Dom Tree pattern as an example to demonstrate how the DOM Tree policies help us to reason comfortably about the preservation of properties of nodes in the presence of unknown code execution. A motivation for using the DOM Tree pattern is how guest/unknown code (such as third party scripted advertising or widgets) are becoming more pervasive in the web today (where details on reasoning about them within the object-capability framework can be found in the work by Maffeis et al.\cite{maffeis2010}). Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years, and we recognise that further work has to be done in the field of smart contracts on the blockchain technology.

\subsection{Overview of Paper}
We first give an overview background history on the OCap model in \cref{sec:ocapmodel}. Readers familiar with the object-capability model and its languages are encouraged to skim through this section. In \cref{sec:specs}, we formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other. Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language. That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapreasoning}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily. We further solidify how our methodology can be used to reason convincingly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}. We end the paper with some preliminary insights on Ethereum in \cref{sec:ethereum}.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{Object-Capability Model}\label{sec:ocapmodel}
\subsection{From Object to Capability}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}. When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.

@@

%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\newpage
\section{Formal specifications}\label{sec:specs}
In this section we describe a small object-oriented language in \cref{sec:speclang} and define predicates that allow us to specify object-capability policies in \cref{sec:permauth}. For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}. To express more precisely how an object can interact with another object, we give formal definitions to the ideas of permission and authority which were studied but not formally defined in \cite{miller2006}. They were later studied in several ways, e.g. in the work by Maffeis et al.\cite{maffeis2010}. Our formal definition is based on but not identical to that from  the work by Drossopoulou et al.'s\cite{drossopoulou2016}. In their papers and our paper, we generally try to capture the intuition of:
\begin{itemize}
\item permission: access right of an object to invoke a behaviour of an object
\item authority: ability to cause effects
\end{itemize}
In this work however, we contribute to the literature by taking a novel approach in formally defining both permission and authority. For permission, we have four different modes permissions specified by four MayAccess predicates. Given a module M and state \losigma,:
\begin{tabbing}
MayAccess$^{Dir,Now}$(o,o$'$): \=o has an object reference of o$'$\\
MayAccess$^{Ind,Now}$(o,o$'$): o contains a reference that points to an object that has a reference\\
\>to another object---thus forming a series of object reference(s)---\\ \> where the last object in the series has an object reference of o$'$\\
MayAccess$^{Dir,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(o,o$'$) holds\\
MayAccess$^{Ind,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(o,o$'$) holds\\
\end{tabbing}
For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}. While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker ---we mean the ability to invoke a method of an object. Therefore implicitly, we consider method invocation as an effect in the system. Given a module M and state \losigma,:
\begin{tabbing}
MayCall(o,o$'$): \=there exists a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
\end{tabbing}

We further elaborate these predicates in \cref{MayAccess} and \cref{MayCall}.
\subsection{Specification Language}\label{sec:speclang}
In our small specification language, we briefly describe the definitions of module, class, runtime state, interpretation, execution, arising configurations, and assertion. All definitions in this section are repeated from the appendix of \cite{drossopoulou2015b} for the reader's convenience. Some parts from \cite{drossopoulou2015b} have also been included in the appendix of this paper. \\

\textbf{Module:}\\
A module maps class identifiers to class descriptions:\\
\begin{logic}
M \loin\ Module = ClassId \locup\ SpecId\\\ablock\qquad\qquad\quad $\rightarrow$ \\
\ablock \qquad \qquad \quad (ClassDescr \locup\ Specification)
\end{logic}\\

\textbf{Class:}\\
Class definitions describe how a class contributes to the runtime behaviour of a program for which we use its field and method declarations. We define method bodies as consisting a sequences of statements; these statements can be field read or internal field assignments (only allowed if the object is \texttt{this}), conditionals, and method calls.\\

@@

\begin{logic}
\textit{Runtime State definition}\\
\losigma\ \loin\ state \quad\= = frame \lotimes\ heap \\
\lophi \loin\ frame \>= StackId $\rightarrow$ val \\
\lochi \loin\ heap \>= addr $\rightarrow$ object \\
v \loin\ val \>= \{\texttt{null}, \texttt{true}, \texttt{false}\} \locup\ \textit{addr} \locup\ $\mathbb{N}$ \\
StackId \>= \{\textit{this}\} \locup\ VarId \locup\ ParId \\
object \>= ClassId \lotimes (FieldId $\rightarrow$ val)\\
$\iota$,$\iota$',.. \>\loin\ \textit{addr}\\
\end{logic}

@@

Further details can be found in appendix \cref{app:execution}.\\

@@

The set Reach(M,\losigma,stmts) collects all configurations reachable during execution of \losigma,stmts. Note that the function Reach(M,\losigma,stmts) is defined even when execution diverges so it may be an infinite set. We therefore can give meaning to capability policies without requiring termination.\\
The definition of Reach can be found in appendix \cref{app:reach}.\\

@@

We say that a class \sf{C} satisfies an assertion \sf{A} in the context of a module \sf{M}, if in all runtime configuration \losigma$'$ arising from execution of any module \sf{M$'$} linked with module \sf{M}, all objects of class \sf{C} satisfy \sf{A}:\\

@@

M,\losigma\ \loturns\ \=MayAccess$^{Dir,Now}$(x,e) \loiff \\
\>\loexists f. \loexec{x.f}{\losigma} = \loexec{e}{\losigma}
\loor\
\loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y}{\losigma} = \loexec{e}{\losigma}\\
\\
 M,\losigma\ \loturns\ \=MayAccess$^{Dir,Eve}$(x,e) \loiff \\
\>\loexists \losigma$'$ \loin Arising(M,\losigma).
M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(x,e)\\
\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Now}$(x,e) \loiff \\
\>\loexists \loconj{f}. \loexec{x.\loconj{f}}{\losigma} = \loexec{e}{\losigma}
\loor\
\loexec{\texttt{this}}{\losigma} = \loexec{x}{\losigma} \loand\ \loexists y. \loexec{y.\loconj{f}}{\losigma} = \loexec{e}{\losigma}\\
\\
M,\losigma\ \loturns\ \=MayAccess$^{Ind,Eve}$(x,e) \loiff \\
\>\loexists \losigma$'$ \loin Arising(M,\losigma). M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(x,e)\\
\hr
\end{logic}

@@

Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions. Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions. \\
\begin{table}[htb]
\caption{MayAccess Relations - Observations}\label{tab:accessrelations}
\centering
\begin{tabular*}{0.5\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &\\
\end{tabular*}
\end{table}\\
We surmarise the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}. Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities. MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$). The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).

@@

\begin{logic}
\hr\\
\emp{Definition---\emp{[MayAffect from Drossopoulou et al.\cite{drossopoulou2016}]}}\\
M,\losigma\ \loturns\ \=MayAffect(x,e) if there exists a method \sf{m}, argument(s) \loconj{a}, state \losigma', such that:\\
\ablock M,\losigma\ \=e.m(\loconj{a}) \loleadsto\ \losigma', and \loexec{e}{M,\losigma} \loneq\ \loexec{e$'$}{M,\losigma$'$}\\
\hr
\end{logic}

@@

Our MayCall(o,o$'$) predicate is not concerned whether or not the state of o$'$ is modified ---the fact that a method in o$'$ can be eventually invoked by o is good enough grounds for us to say that o has authority of o$'$. Our justification is that in the object-oriented paradigm, programmers almost always enforce security of a sensitive field in a protected object to be declared private, and define some public method in the object that can modify the concerned private field based on some condition. Hence, the question of whether an \textit{untrusted} object can modify a private field in an object \sf{o$'$} can be expressed as a question of whether the untrusted object can invoke a method of \sf{o$'$}. Given our definitions of MayCall(o,o$'$), it follows that a necessary condition for modification of a private field in o$'$ is the invocation of a method of o$'$. Proving the falsity of our MayCall(o,o$'$) predicate is therefore sufficient to prove that for a field \sf{f} in o$'$, MayAffect(o,o$'$.f) does not hold.\\
\begin{logic}
\hr\\
\emp{Lemma---[Field Modification Requires Authority]}\\
\loforall \losigma$'$,f. [\=\losigma$'$\loin Arising(M,\losigma) \loand\ \loexec{o$'$.f}{\losigma$'$}\loneq\loexec{o$'$.f}{\losigma}\loimplies \loexists o. M,\losigma\ \loturns\ MayCall(o,o$'$) ]\\
\hr
\end{logic}
